Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-15 IEEE_ConferenceOnGames/new_article.tex
new document
more approach

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote.
% If that is unneeded, please comment it out.
\usepackage[style=ieee,backend=biber,style=numeric,sorting=none]{biblatex}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\addbibresource{References/article_references.bib}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

\section{State of the art}
\label{SoA}
In this Section we present PCG works relevant to the Level generation for the
Angry Birds video game. Renz et al. \cite{Renz, Renz2015TheAB} organized an AI
competition with the aim of developing intelligent agents that could play new
Angry Birds levels better than the best human players. The competition was then
collocated in other AI conferences. In order to train and test these agents,
more novel and some times difficult levels must be generated. A related
competition was then proposed by Stephenson et al. \cite{Stephenson} this time
with the goal of developing a level generator for the game. Several generators
have been proposed in the following years of competition. M. Kaidan et al.
\cite{Kaidan2015} propose a model in which the generation of content is
controlled by the skill of a player while actively playing the game, the moment
the player finishes a certain level the algorithm obtains the score and
generates the next level to be played. Y. Jiang et al. \cite{Jiang2017} proposed
the use of predefined letter style patterns and combined to create a small pool
of words and letters used to be combined in order to create levels with text on
them, the participants would play the game and after that they would be asked if
they liked the layout of the levels. A search based approach has been also used
in several works, recently a

\section{Approach}

In this Section we describe the approaches that were used for this project, first 
explaining some information of how techniques like OOE were implemented with the paper.

The game contains a total of 11 pieces as shown on figure \ref{piece_list} that
cannot be modified, it is not possible to add more or modify the existing ones,
with this in mind the purpose is to create composites of this same pieces in
different locations and angles in order to create structures that can be used as
building blocks for more complex levels.

Since there is no way to \textit{add} a composite to the pool of pieces of the 
game then a way to combine both groups is needed, in order to do this first the 
original 11 pieces of the game will be converted into composites, since the easiest
way to read the data to create composites is by using lists then a \textit{composite} 
of one element will be created for each piece, after this in order to add a 
composite of pieces as a new component first a group of existing composites is 
selected and placed in a imaginary enviroment then as a unique group they are 
measured, an imaginary bounding box is defined by calculating the lengths of all
the pieces, finding the borders of said pieces and then finding the points that could
define a box around all of them, then use this defined box to calculate the 
height and width of the new composite and then calculate from 
the center point of said box the offsets from the center of the
group to the center of each respective piece as shown on figure
\ref{bounding_boc_calc}, this information is needed in order to add it as a new
composite so it can be used to place the pieces in their respective new positions
when the files needed for the simulation are being created.

So using the composite generator the order of this sequence of
events needs to be done as follows:

@@

\begin{enumerate}
\item Gnerate composited to add to the pool of options.
\item Generate the first generation of the population.
\item Run a Simulation.
\item Evaluate each member of the population.
\item If a member of the population has remaining pieces obtain the location and
angle of the pieces, create a composite and add it to the pool.
\item Select the parents for the next generation.
\item Do the crossover and mutation operations.
\item Repeat the cycle.
\end{enumerate}

@@

The Angry Birds level generation track is a competition where the objective is 
to generate levels that are \textit{fun} and \textit{enjoyable} however they 
also have to be challenging but imposible however the rules for the competition 
specify that a configuration file must be used to
generate the levels of the game in a limited amount of time, the information on
the configuration file is first the number of \textbf{\textit{different}} levels
that must be generated, the second value is the different materials \textit{not
allowed} to be used, for instance a group of generated levels must not have
circles which composition material is ice or stone moreover it does not have to
integrate a triangle piece with ice, stone or wood material so this combinations
must be removed from the pool of options, the third configuration value is the
\textit{number of pigs} that must be placed in a level, and finally the
\textit{limit} of time that the system has to create the required amount of
levels.

@@

\subsection{Original approach}
% This sentence needs more work -MG
During the original phase of the design of the system, it was decided that the way 
the composites were going to evolve was by first placing only base pieces on a level 
then after a simulation the remaining pieces
of a member of the population were going to be used has a new composite if after a 
simulation a group was still standing and grouped on a single point, their individual 
offsets from the center of the group was calculated and then using this information a new 
composite could be added to the pool of options, this new
% we don't use member of the populations, use proposed solution, candidate solution
% or individual of the population. - MG  
additions to the composite pool could then be added to the population by the mutation
operation when a new member of the population was being created, the objective 
in this first phase of the project was to generate
structures that were capable of maintaining their own balance as shown on figure
\ref{test_old}, since the objective of open-ended evolution for this project is
to add new composites to the pool of options, then the algorithm will first run a
simulation with a group of pieces and after it the ones that were not destroyed
by falling because of gravity will be saved and added to the pool of options.

@@

Since there has to be a way to represent the struccture of a level in a a way that
both the algorithm as well as the game are able to interpret it \cite{togelius2016Representationsforsearch-basedmethods}, 
then the way to represent the elements of a member of the population in thos work 
is by using the id of each piece and the new added ones has a pointer to the respective data
that needs to be obtained, this way and individual can be represented as shown
on figure \ref{old_chrom} where each chromosome represents a certain item in the
pool of pieces, the numbers for this elements are assigned in a first come first
served basis, the numbers are infinitely incremented and the same pieces can be
repeated on each individual of the population.

@@

\printbibliography
\newrefcontext[sorting=ydnt]

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

    \section{State of the art}
    \label{SoA}
    In this Section we present PCG works relevant to the Level generation for the
    Angry Birds video game. Renz et al. \cite{Renz, Renz2015TheAB} organized an AI
    competition with the aim of developing intelligent agents that could play new
    Angry Birds levels better than the best human players. The competition was then
    collocated in other AI conferences. In order to train and test these agents,
    more novel and some times difficult levels must be generated. A related
    competition was then proposed by Stephenson et al. \cite{Stephenson} this time
    with the goal of developing a level generator for the game. Several generators
    have been proposed in the following years of competition. M. Kaidan et al.
    \cite{Kaidan2015} propose a model in which the generation of content is
    controlled by the skill of a player while actively playing the game, the moment
    the player finishes a certain level the algorithm obtains the score and
    generates the next level to be played. Y. Jiang et al. \cite{Jiang2017} proposed
    the use of predefined letter style patterns and combined to create a small pool
    of words and letters used to be combined in order to create levels with text on
    them, the participants would play the game and after that they would be asked if
    they liked the layout of the levels. A search based approach has been also used
    in several works. 
    %%%
    % TO DO: Laura´s work here 
    %%%

    \section{Proposed Method}

    In this Section we describe the approach used in this work, explaining first,
    the restrictions of the game.
    The game contains a total of 11 block shapes as shown in Figure \ref{piece_list}.
    These basic blocks cannot be modified, it is not possible to add more or modify the existing ones,
    with this in mind the purpose is to create composites of these blocks in
    different locations and angles in order to create structures that can be used as
    building blocks for more complex structures.

    Since there is no way to \textit{add} a new composite to the pool of blocks,
    then a way to combine both is needed, in order to do this first the original
    11 pieces of the game will be converted into composites of a single element.
    A composite is defined as a lists of blocks, then a \textit{composite} of
    one element will be created for each piece. Composites are the first layer
    the building blocks for the structures to come. A composite could include
    different types of blocks, rotaded at a certain angle and compose a basic
    structure. An ideal composite could support other composites on top with out
    falling apart, but this is not required. A composite is measured using
    maginary bounding box, defined by calculating the lengths of all the pieces,
    finding the borders of said pieces and then finding the points that could
    define a box around all of them, then use this defined box to calculate the
    height and width of the new composite and then calculate from the center
    point of said box the offsets from the center of the group to the center of
    each respective piece as shown on figure \ref{bounding_boc_calc}, this
    information is needed in order to add it as a new composite so it can be
    used to place the pieces in their respective new positions when the files
    needed for the simulation are being created. This is similar to creating a
    group of elements in a drawing application.

    %% 
    Good composites can be found using an evolutionary approach proposed earlier 
    by Calle and Merelo, where basic structures are generated by searching a space
    of falling blocks and selecting those that remain standing. This algorithm is 
    not capable of creating complex strucures, but it can generate very novel basic
    structures. We can see that the evolution at this level can benefit from having
    more freedom in how basic structures are built. We do not want to form piles
    or sequences of blocks, this will be done at a higher level in the composite hirarchy.

    At his level the composite generator has the following steps, for more details refer to the paper:

@@

    \begin{enumerate}
    \item Generate the first generation of the population, consisting of randomly generated list of blocks.
    \item Run a Simulation.
    \item Evaluate each individual of the population.
    \item If a member of the population has remaining pieces obtain the location and
    angle of the pieces, assign a fitness.
    \item Do the selection, crossover and mutation operations.
    \item Repeat the cycle.
    \item From the best individuals select composites. Several algorithms could be proposed for this step, for instance selcting clusters of blocks.
    \end{enumerate}

    Another option is to just manually generate interesting composites to be used by the algorithm.
    We could say that the previous approach is an option for a procedural generation of composites.

@@

    The Angry Birds level generation track is a competition where the objective is 
    to generate levels that are \textit{fun} and \textit{enjoyable} however they 
    also have to be challenging but not imposible. However the rules for the competition 
    specify that a configuration file must be used to
    generate the levels of the game in a limited amount of time, the information on
    the configuration file is first the number of \textbf{\textit{different}} levels
    that must be generated, the second value is the different materials \textit{not
    allowed} to be used, for instance a group of generated levels must not have
    circles which composition material is ice or stone, moreover it does not have to
    integrate a triangle piece with ice, stone or wood material so this combinations
    must be removed from the pool of options, the third configuration value is the
    \textit{number of pigs} that must be placed in a level, and finally the
    \textit{limit} of time that the system has to create the required amount of
    levels.

@@

    \subsection{Original approach}
    % This sentence needs more work -MG

    During the original phase of the design of the system, it was decided that the way 
    composites were going to evolve was by first placing only base pieces on a level 
    then after a simulation the remaining pieces
    of a member of the population were going to be used as a new composite if after a 
    simulation a group was still standing and grouped on a single point, their individual 
    offsets from the center of the group was calculated and then using this information a new 
    composite could be added to the pool of options, this new
    % we don't use member of the populations, use proposed solution, candidate solution
    % or individual of the population. - MG  
    additions to the composite pool could then be added to the population by the mutation
    operation when a new member of the population was being created, the objective 
    in this first phase of the project was to generate
    structures that were capable of maintaining their own balance as shown on figure
    \ref{test_old}, since the objective of open-ended evolution for this project is
    to add new composites to the pool of options, then the algorithm will first run a
    simulation with a group of pieces and after it the ones that were not destroyed
    by falling because of gravity will be saved and added to the pool of options.

@@

    Since there has to be a way to represent the structure of a level in a a way that
    both the algorithm as well as the game are able to interpret it \cite{togelius2016Representationsforsearch-basedmethods}, 
    then the way to represent the elements of a member of the population in thos work 
    is by using the id of each piece and the new added ones has a pointer to the respective data
    that needs to be obtained, this way and individual can be represented as shown
    on figure \ref{old_chrom} where each chromosome represents a certain item in the
    pool of pieces, the numbers for this elements are assigned in a first come first
    served basis, the numbers are infinitely incremented and the same pieces can be
    repeated on each individual of the population.

@@

\bibliographystyle{IEEEtran}
\bibliography{References/article_references.bib}
